---
title: "Reproducible Research Week 2 Course Project 1"
author: "Irfan Ahmed"
date: "5/20/2020"
output: html_document
---


Loading and preprocessing the data

```{r}activity <- read.csv("~/DataScience/Rep Research/RepData_PeerAssessment1-master/activity.csv")
```
Adding total steps daily ignoring NAs

```{r}dailysteps <- activity %>% 
                group_by(date) %>% 
                summarize(sumsteps = sum(steps, na.rm=TRUE))
```

Histogram of the total number of steps taken each day

```{r}hist(dailysteps$sumsteps, col ="red", xlab = "Daily Steps", 
     main = "Histogram of total number of steps each day", ylim = c(0,31))
```

Mean and median number of steps taken each day
```{r}meansteps <- mean(dailysteps$sumsteps)
mediansteps <- median(dailysteps$sumsteps)
paste("The mean is:", meansteps)
paste("The median is:", mediansteps)
```
Time series plot of the average number of steps taken
```{r}stepsPerInterval <- activity %>%
    group_by(interval) %>%
    summarize(meansteps = mean(steps, na.rm = TRUE))
```
```{r}plot(stepsPerInterval$meansteps ~ stepsPerInterval$interval,
    col="red", type="l", xlab = "5 Minute Intervals", ylab = "Average Number of Steps",
    main = "Steps By Time Interval")
```
The 5-minute interval that, on average, contains the maximum number of steps
```{r}print(paste("Interval where the maximum number of steps is: ", stepsPerInterval$interval[which.max(stepsPerInterval$meansteps)]))
```
```print(paste("Average steps for that interval: ",round(max(stepsPerInterval$meansteps), digits = 2)))
```

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}print(paste("Total number of NAs are: ", sum(is.na(activity$steps))))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}activityNoNA <- activity  
for (i in 1:nrow(activity)){
    if(is.na(activity$steps[i])){
        activityNoNA$steps[i]<- stepsPerInterval$meansteps[activityNoNA$interval[i] == stepsPerInterval$interval]
    }
}
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}dailystepsNoNA <- activityNoNA %>%
         group_by(date) %>%
         summarize(sumsteps = sum(steps, na.rm = TRUE))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}hist(dailystepsNoNA$sumsteps, col ="red", xlab = "Daily Steps", main = "Histogram of total number of steps each day", ylim = c(0,31))
```

```{r}meanPostNA <- round(mean(stepsPerDay$sumsteps), digits = 2)
medianPostNA <- round(median(stepsPerDay$sumsteps), digits = 2)
```
```{r}print(paste("The mean is: ", mean(meanPostNA)))
print(paste("The median is: ", median(medianPostNA)))
```
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}activityDoW <- activityNoNA
activityDoW$date <- as.date(activityDoW$date)
activityDoW$day <- ifelse(weekdays(activityDoW$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
activityDoW$day <- as.factor(activityDoW$day)
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}activityWeekday <- filter(activityDoW, activityDoW$day == "weekday")
activityWeekend <- filter(activityDoW, activityDoW$day == "weekend")

activityWeekday <- activityWeekday %>%
    group_by(interval) %>%
    summarize(steps = mean(steps)) 
activityWeekday$day <- "weekday"

activityWeekend <- activityWeekend %>%
    group_by(interval) %>%
    summarize(steps = mean(steps)) 
activityWeekend$day <- "weekend"

wkdayWkend <- rbind(activityWeekday, activityWeekend)
wkdayWkend$day <- as.factor(wkdayWkend$day)
```

```{r}g <- ggplot (wkdayWkend, aes (interval, steps))
g + geom_line() + facet_grid (day~.) + 
    theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14)) + 
    labs(y = "Number of Steps") + labs(x = "Interval") + 
    ggtitle("Average Number of Steps - Weekday vs. Weekend") + 
    theme(plot.title = element_text(hjust = 0.5))
```